### 2019-04-02 5_evaluation.tex
Apply Alex's and George's comments.

We present a comprehensive evaluation of Task Bench on up to 256
Haswell nodes of the Cori supercomputer~\cite{Cori}, a Cray XC40
machine. Cori Haswell nodes have 2 sockets with Intel Xeon E5-2698 v3
processors (a total of 32 physical cores per node), 128 GB RAM, and a
Cray Aries interconnect. We use GCC 7.3.0 for compiling all Task Bench
implementations, and (where applicable) the system default MPI
implementation, Cray MPICH 7.7.3. Additional versions and flags for the
various systems are shown in Table~\ref{tab:flags}.

@@

@@

Figure~\ref{fig:flops} shows the FLOPS achieved with a compute-bound
kernel with varying problem sizes. This is the full version of
Figure~\ref{fig:flops-mpi}. In the best case, we measure peak FLOPS of
$1.26 \times 10^{12}$, which compares favorably with the officially
reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
of measuring efficiency, we use our empirically determined number as
the baseline for 100\% efficiency.

@@

In most cases, task granularity has plateaued prior to the 50\% efficiency point,
though some systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold compared to
measuring task scheduling throughput in the presence of empty tasks
(effectively METG(0\%)). Empty tasks permit runtime
systems to make unrealistic tradeoffs which are not applicable to real
application workloads where some degree of efficiency is necessarily a
goal. Reporting numbers with empty tasks therefore gives unfair
advantage to systems that make such unrealistic tradeoffs.

@@

One of the basic questions when considering different programming
systems is how much overhead the system introduces. This can be a
tricky question to answer directly because some systems introduce
overhead \emph{inline} (i.e., by running system internal processes on
the same cores as application tasks), while other systems introduce
overhead \emph{out-of-line} (i.e., by dedicating one or more cores
solely to runtime use). Some systems, like Charm++, Realm, and Regent,
even support both configurations.

To answer this question, we use the METG metric to determine the
smallest task granularity which can be executed at a given level of
efficiency, as a proxy for overhead. Figure~\ref{fig:metg-compute}
shows how METG(50\%) varies with node count for a variety of
dependence patterns supported by Task Bench. METG(50\%) is calculated
separately at each node count, to isolate runtime system behavior from
changes in communication latencies and topology when using
progressively larger portions of the machine. We consider the
following configurations of Task Bench:

@@

We observe in the results that the baseline overheads of different
systems can vary by over 5 orders of magnitude. It is worth
remembering when considering this metric that this is a \emph{minimum}
effective task granularity. Therefore applications with an average
task granularity of \emph{at least} this value can usually be expected
to execute efficiently. Typical average task granularity will
generally be determined by the application domain being
considered. Most notably, for data analytics workloads being run in
industrial data centers, the higher METG values observed for Spark and
TensorFlow are sufficient. In contrast, for high-performance
scientific simulations, task granularities in the millisecond range
are useful, as such applications communicate (e.g., for halo
exchanges) much more frequently.

@@

METG is a useful metric in part because it summarizes the basic task
scheduling overhead of a system in a single number. This makes it
possible to evaluate METG at different node counts (shown in
Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
communication topology and latency.

In general, the least complicated pattern (stencil) is most favorable
to MPI, as it represents a truly bulk synchronous task graph with no
opportunity for task parallelism. For the stencil pattern, the
dominating factor is the basic overhead of executing a task, which is
minimal for MPI as the Task Bench implementation is bulk synchronous
and simply executes tasks one after another in alternation with
communication phases. In contrast the asynchronous execution
mechanisms of the other systems are pure overhead in this scenario.

The gap between MPI and other systems shrinks as the complexity of the
communication pattern grows, and even reverses as task parallelism is
added in the form of multiple task graphs.

@@

Implicitly parallel systems such as PaRSEC, StarPU and Regent that
rely on runtime analysis to build the DAG can suffer from
scalability bottlenecks if every node must consider the tasks
executing on all other nodes. PaRSEC DTD and StarPU mitigate this
partially by allowing the user to omit tasks that are not direct
dependencies of those executed on the current node. However, this DAG
trimming approach still requires some dynamic checks that scale with
the number of nodes and thus limits scalability. Approaches based on compile-time analysis can
partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
still retains dynamic checks leading to limited scalability. Regent
uses a compile-time optimization that enables full horizontal
scalability~\cite{ControlReplication17}. A similar, manual
optimization is shown in the plots as PaRSEC Shard.

@@

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.3\times$ to $59\times$ (median $3.0\times$). From 1 to 3
dependencies, the ratio varies from $1.2\times$ to $15\times$ (median
$2.1\times$). The difference is most pronounced in systems which
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 us with 3 dependencies,
a factor of $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels
as quickly as possible.

@@

One advantage of systems with asynchronous execution capabilities is
the ability to mitigate load imbalance, especially in the presence of
task parallelism. To quantify this effect,
Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
efficiency curves under load imbalance with uniformly distributed task
durations. The task durations are generated with a deterministic
pseudo random number generator with a deterministic seed to ensure
that the task durations under all systems are identical.

@@

The difference between the remaining systems is due primarily to the
scheduling behavior of the various systems. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Therefore, systems which provide an
additional on-node work stealing capability see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at very small task
granularities. This can be seen in the difference between Chapel
Distrib (which uses a work stealing scheduler) vs Chapel with its
default (non-work-stealing) scheduler.

@@

Asynchronous systems such as Charm++ are able to show two benefits in
these plots. First, by overlapping communication with computation,
such systems are able to execute smaller task granularities at higher
levels of efficiency compared to the bulk synchronous MPI
implementation. Second, the asynchrony and scheduling flexibility from
executing multiple graphs also makes the curves smoother,
as spikes in latency due to interference from other jobs can be
mitigated, leading to more predictable performance, especially at
smaller message sizes.

>>>>

We present a comprehensive evaluation of our Task Bench implementations on up to 256
Haswell nodes of the Cori supercomputer~\cite{Cori}, a Cray XC40
machine. Cori Haswell nodes have 2 sockets with Intel Xeon E5-2698 v3
processors (a total of 32 physical cores per node), 128 GB RAM, and a
Cray Aries interconnect. We use GCC 7.3.0 for compiling all Task Bench
implementations, and (where applicable) the system default MPI
implementation, Cray MPICH 7.7.3. Additional versions and flags for the
various systems are shown in Table~\ref{tab:flags}.

@@

Figure~\ref{fig:flops} shows the FLOPS achieved with a compute-bound
kernel with varying problem sizes. This is the full version of
Figure~\ref{fig:flops-mpi}. Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern. In the best case, we measure peak FLOPS of
$1.26 \times 10^{12}$, which compares favorably with the officially
reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
of measuring efficiency, we use our empirically determined number as
the baseline for 100\% efficiency.

@@

@@

In most cases, task granularity asymptotes prior to the 50\% efficiency point,
though some systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold instead of
measuring task scheduling throughput of empty tasks
(effectively METG(0\%)). Measuring performance using empty tasks can
reward implementation strategies, such as devoting nearly 100\% of
system resources to the runtime system, that make no sense for real
applications.

@@

One of the basic questions when considering different programming
systems is: How much overhead does the system add? This question is tricky to answer directly because some systems introduce
overhead \emph{inline} (i.e., by running system internal processes on
the same cores as application tasks), while other systems introduce
overhead \emph{out-of-line} (i.e., by dedicating one or more cores
solely to runtime use). Some systems, like Charm++, Realm, and Regent,
even support both configurations.

To answer this question, we use the METG metric to determine the
smallest task granularity that can be executed at a given level of
efficiency as a proxy for overhead. Figure~\ref{fig:metg-compute}
shows how METG(50\%) varies with node count for a variety of
dependence patterns supported by Task Bench. METG(50\%) is calculated
separately at each node count, to isolate runtime system behavior from
changes in communication latencies and topology when using
progressively larger portions of the machine. We consider the
following configurations of Task Bench:

@@

We observe in the results that the baseline overheads of different
systems vary by over 5 orders of magnitude. It is worth
remembering when considering this metric that this is a \emph{minimum}
effective task granularity. Therefore applications with an average
task granularity of \emph{at least} this value can usually be expected
to execute efficiently. Typical average task granularity will
generally be determined by the application domain being
considered. Most notably, for data analytics workloads being run in
industrial data centers, the higher METG values observed for Spark and
TensorFlow are sufficient. In contrast, for high-performance
scientific simulations, task granularities in the millisecond range
are useful, as such applications communicate (e.g., for halo
exchanges) much more frequently.

In general, the least complicated pattern (stencil) is most favorable
to MPI, as it represents a truly bulk synchronous task graph with no
opportunity for task parallelism. For the stencil pattern, the
dominating factor is the basic overhead of executing a task, which is
minimal for MPI as the Task Bench implementation is bulk synchronous
and simply executes tasks one after another in alternation with
communication phases. In contrast, the asynchronous execution
mechanisms of the other systems are pure overhead in this scenario.

The gap between MPI and other systems shrinks as the complexity of the
communication pattern grows, and even reverses as task parallelism is
added in the form of multiple task graphs. Some of the slower systems
are omitted from comparisons with more complicated dependence
patterns, as they failed to complete in a reasonable time.

@@

METG is a useful metric in part because it summarizes the basic task
scheduling overhead of each programming system in a single number. This makes it
possible to evaluate METG at different node counts (shown in
Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
communication topology and latency.

@@

Implicitly parallel systems such as PaRSEC, StarPU and Regent that
rely on runtime analysis to build the DAG can suffer from
scalability bottlenecks if every node must consider the tasks
executing on all other nodes. PaRSEC DTD and StarPU mitigate this
partially by allowing the user to omit tasks that are not direct
dependencies of those executed on the current node. However, this DAG
trimming approach still requires some dynamic checks that scale with
the number of nodes and thus limits scalability. Approaches based on compile-time analysis can
partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
still retains dynamic checks leading to limited scalability. Regent
uses a compile-time optimization that enables full horizontal
scalability~\cite{ControlReplication17}. A similar, manual
optimization is shown in the plots as PaRSEC DTD Shard.

@@

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.3\times$ to $59\times$ (median $3.0\times$). From 1 to 3
dependencies, the ratio varies from $1.2\times$ to $15\times$ (median
$2.1\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 us with 3 dependencies,
a factor of $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels
as quickly as possible.

@@

One advantage of systems with asynchronous execution capabilities is
the ability to mitigate load imbalance, especially in the presence of
task parallelism. To quantify this effect,
Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
efficiency curves under load imbalance with uniformly distributed task
durations. The task durations are generated with a deterministic
pseudo random number generator with a deterministic seed to ensure
identical task durations for all systems.

@@

The difference between the remaining systems is due primarily to the
scheduling behavior of the various systems. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Therefore, systems that provide an
additional on-node work stealing capability see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at very small task
granularities. This can be seen in the difference between Chapel
distrib (which uses a work stealing scheduler) vs Chapel with its
default (non-work-stealing) scheduler.

@@

Asynchronous systems such as Charm++ are able to show two benefits in
these plots. First, by overlapping communication with computation,
such systems execute smaller task granularities at higher
levels of efficiency compared to the bulk synchronous MPI
implementation. Second, the asynchrony and scheduling flexibility from
executing multiple graphs also makes the curves smoother,
as spikes in latency due to interference from other jobs can be
mitigated, leading to more predictable performance, especially at
smaller message sizes.

### 2019-04-02 7_conclusion.tex
Apply Alex's and George's comments.

Parameterized benchmarks are a promising new route for evaluating the
performance of parallel and distribute programming systems. By
composing a small number of orthogonal primitives, parameterized
benchmarks make it possible to explore a space of application
scenarios, while minimizing the marginal cost of developing each
additional feature. In this case of Task Bench, this has enabled a
comprehensive comparative study of programming system performance
which is unique in both the breadth of systems considered and the
depth of insight achieved about them.

@@

Not considered in our analysis is the impact of programming system
features on programmer productivity and performance portability. Most
applications do not operate at the absolute extreme of runtime-limited
performance, and thus may choose to trade overhead for better
usability. Our study helps to quantify exactly how much a tradeoff is
being made, so that users can be better informed, and developers can
see the impact of features of the performance of their programming
systems.

>>>>

Parameterized benchmarks are a promising new route for evaluating the
performance of parallel and distributed programming systems. By
composing a small number of orthogonal primitives, parameterized
benchmarks make it possible to explore a space of application
scenarios, while minimizing the marginal cost of developing each
additional feature. In the case of Task Bench, this has enabled a
comprehensive comparative study of programming system performance
which is unique in both the breadth of systems considered and the
depth of insight achieved about them.

@@

Not considered in our analysis is the impact of programming system
features on programmer productivity and performance portability. Most
applications do not operate at the absolute extreme of runtime-limited
performance, and thus may choose to trade overhead for better
usability. Our study helps to quantify the performance side of that tradeoff so that users can be better informed and developers can
see the impact of features on the performance of their programming
systems.
